Inferring Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-10-01 doc/paper-lcr.tex
draft update

The time-delay distance is defined as
\be \label{eq:dt}
\Dt \equiv (1+\zd) D_{\rm d} D_{\rm s}/D_{\rm ds},
\ee
with which Equation \ref{eq:T} can be rephrased as
\be
\Delta t(\bmath{\theta},\bmath{\beta})  =  \frac{\Dt}{c} \, \phi(\bmath{\theta},\bmath{\beta})    ,
\ee 
Since the time delays are proportional to $1/ (1+ \kappax)$, if there is any external convergence, that is not included in the lens
modeling, then the time delay distance -- inferred assuming $\kappax = 0$ -- will be
$(1-\kappax)$ less than the true value of $\Dt$.
\be 
\label{eq:MassSheet:H0bias}
\Dt^{\rm{true}}=(1-\kappax) \Dt^{{\kappax = 0}}.
\ee
We can overcome this degeneracy if we have additional knowledge of $\kappax$.

@@

In this work our aim is to quantify the lensing effect of mass-sheets that are not physically associated with the lens galaxy -- mass-sheets
lying infront or behind the lens -- and the uncertainties induced by reconstructing
those mass-sheets using simple prescriptions applied to potential astronomical observables.
We also aim to quantify how uncertainties in the
estimations of $\kappax$ propogate into uncertainties on the cosmological parametres
which can be inferred from measurements of $\Dt$.

@@

For real lines of sight, 
\phil{we would like to estimate the external convergence as best we can given all
data available. Typically this data is in the form of a photometric galaxy
catalog, containing positions, brightnesses, colours, and other type
information for each galaxy in the field of the lens, down to some magnitude
limit. Since we do not observe halo mass directy, we need some way estimating
the mass of each galaxy in the catalog, so that we can compute its
contribution to $\kappax$. This mass assignment recipe will be uncertain, but
can be calibrated to cosmological simulations.}

@@

Unphysically, if one integrates the density profile of an NFW profile, the total mass is divergent.
Similarly, if the universe is homogeneously populated with NFW halos, the projected surface mass along any
line of sight will also be divergant\footnote{At large radius, $\Sigma_{\rm
nfw} \propto R^{-2}$, but the differential number of halos centred within an
annulus of width $\dee R$ is given by $\dee N_{\rm annulus} \propto R \dee R$,
so  $\Sigma_{\rm total} \propto \int_{0}^{\infty} R^{-1} \dee R$, which
diverges logarithmically}. Since infinite mass is unphysical, the profile must
be truncated at some point. \todo{Tom}{Add notes on previous truncation results}. Several truncation profiles have been suggested \citep[e.g][]{BMO}, but beyond several virial radii, the amount of matter associated with a halo is likely to be low; \citet{Nandra2012} showed that the Universe's expansion can place an upper limit on the radius of NFW halos -- for cluster scale halos, this happens at around 25 times the NFW scale radius. In \Sref{sec:knownMh+z} we investigate how different truncations affect the recovered total convergence for a line of sight, using the truncated NFW profile of \citet{BMO}
\be\label{eq:bmoprofile}
\rho(r) = 
\frac{\rho_{\rm NFW} (r)}{1+\left(r/r_{t}\right)^2},
\ee
which is the same as the NFW profile in the limit that $r_t$ goes to infinity. The convergence and shear for the BMO profile are given in Appendix \ref{appendix:halos}.

>>>>

The time-delay distance is defined as
\be \label{eq:dt}
\Dt \equiv (1+\zd) D_{\rm d} D_{\rm s}/D_{\rm ds},
\ee
with which Equation \ref{eq:T} can be rephrased as
\be
\Delta t(\bmath{\theta},\bmath{\beta})  =  \frac{\Dt}{c} \, \phi(\bmath{\theta},\bmath{\beta})    ,
\ee 
Since the time delays are proportional to $1/ (1- \kappax)$, if there is any external convergence, that is not included in the lens
modeling, then the time delay distance -- inferred assuming $\kappax = 0$ -- will be
$(1-\kappax)$ less than the true value of $\Dt$.
\be 
\label{eq:MassSheet:H0bias}
\Dt^{\rm{true}}=(1-\kappax) \Dt^{{\kappax = 0}}.
\ee
We can overcome this degeneracy if we have additional knowledge of $\kappax$.

@@

In this work our aim is to quantify the lensing effect of mass-sheets that are not physically associated with the lens galaxy -- mass-sheets
lying infront or behind the lens -- and the uncertainties induced by reconstructing
those mass-sheets using simple prescriptions applied to potential astronomical observables.
\comments{We also aim to quantify how uncertainties in the
estimations of $\kappax$ propogate into uncertainties on the cosmological parametres
which can be inferred from measurements of $\Dt$.}

@@

For real lines of sight, 
\phil{we would like to estimate the external convergence as best we can given all
data available. Typically this data is in the form of a photometric galaxy
catalog, containing positions, brightnesses, colours, and other type
information for each galaxy in the field of the lens, down to some magnitude
limit. Since we do not observe halo mass directy, we need some way of estimating
the mass of each galaxy in the catalog, so that we can compute its
contribution to $\kappax$. This mass assignment recipe will be uncertain, but
can be calibrated to cosmological simulations.}

@@

Unphysically, if one integrates the density profile of an NFW profile, the total mass is divergent.
Similarly, if the universe is homogeneously populated with NFW halos, the projected surface mass along any
line of sight will also be divergant\footnote{At large radius, $\Sigma_{\rm
nfw} \propto R^{-2}$, but the differential number of halos centred within an
annulus of width $\dee R$ is given by $\dee N_{\rm annulus} \propto R \dee R$,
so  $\Sigma_{\rm total} \propto \int_{0}^{\infty} R^{-1} \dee R$, which
diverges logarithmically}. Since infinite mass is unphysical, the profile must
be truncated at some point. \todo{Tom}{Add notes on previous truncation results}. Several truncation profiles have been suggested \citep[e.g][]{BMO}, but beyond several virial radii, the amount of matter associated with a halo is likely to be low; \citet{Nandra2012} showed that the Universe's expansion can place an upper limit on the radius of NFW halos -- for cluster scale halos, this happens at around 25 times the NFW scale radius. In \Sref{sec:knownMh+z} we investigate how different truncations affect the recovered total convergence for a line of sight, using the truncated NFW profile of \citet{BMO}
\be\label{eq:bmoprofile}
\rho(r) = 
\frac{\rho_{\rm NFW} (r)}{1+\left(r/r_{t}\right)^2},
\ee
which is the same as the NFW profile in the limit that the truncation radius, $r_t$ goes to infinity. The convergence and shear for the BMO profile are given in Appendix \ref{appendix:halos}.
